Analysis \& Visualization of Call Center Data using Python

### 2018-09-07 paper.tex
Added, subtracted

The data to be analyzed was acquired in three distinct comma-separated value files.  Two of the files are publicly available, and are also posted on the \href{https://github.com/jdbul33/verbose-chainsaw}{Github repository} for this project.  The third file was received directly from the city of South Bend, and as a result is not posted publicly at this time.
\par
The first file, referred to as the daily data, contains a daily summary of the call center data from the years 2013-2015.  It is organized with 637 records indexed by date, with each record containing information such as the number of calls presented, average wait time, average number of calls in queue, and number of abandoned calls for that day.
\par
The second file, referred to as the case data, contains 488,601 rows of individual call data from 2013 through 2015.  Calls were logged anonymously with data such as date, duration, topic, and department.  This data format went obsolete when the current call system was implemented.
\par
The third file, referred to as the topic data, is the storage method of the current phone system in use by the city.  This file contains 202,500 call records from 2016-2018.  It contains much of the same attributes as the case data, but it is standardized to reflect the city's use of knowledge base articles.  These articles correspond to possible questions or issues raised by callers; as a result, the call center employees have a reference to answer questions.  Additionally, this systemizes the format for recording issues for each call, since each topic has a standardized name.  This ensures topics and departments match across the list, allowing for efficient and accurate analysis.  Since this is the most recent data and is reflective of the city's current systems, this file was heavily relied upon for analysis on a departmental and topical basis.

@@

Exploration of calls by topic was a priority for the city.  Within each department, numerous topics exist.  Given the recent implementation of knowledge base articles (which can be seen in the most recent data), the attachment of topics to each call has been standardized.  This allows for a simplified, more accurate analysis.  Calls by topic were specifically analyzed by average duration and total call volume.
\par
The %left off here

\section{Discussion}

@@

These charts all utilize various implementation of Bokeh tools, such as HoverTool, BoxZoom, Pan, and checkbox interactivity.  By incorporating these interactions into the dashboard, the result is a more dynamic, engaging product that is simple to interact with for employees outside of the technical fields.  An image of the final dashboard can be found at the end of this paper, while a live link can be followed \href{https://jdbul33.github.io/311_Call_Dashboard.html}{here}.

@@

\bibitem{python}
Python Software Foundation. (2016). Python 3.6. https://docs.python.org/3.6/.

\bibitem{anaconda}
Anaconda. (2018). Anaconda Distribution 5.2. https://docs.anaconda.com/anaconda/.

>>>>

The data to be analyzed was acquired in three distinct comma-separated value files.  The first file, referred to as the daily data, contains a daily summary of the call center data from the years 2013-2015.  It is organized with 637 records indexed by date, with each record containing information such as the number of calls presented, average wait time, average number of calls in queue, and number of abandoned calls for that day.
\par
The second file, referred to as the case data, contains 488,601 rows of individual call data from 2013 through 2015.  Calls were logged anonymously with data such as date, duration, topic, and department.  This data format went obsolete when the current call system was implemented.
\par
The third file, referred to as the topic data, is the storage method of the current phone system in use by the city.  This file contains 202,500 call records from 2016-2018.  It contains much of the same attributes as the case data, but it is standardized to reflect the city's use of knowledge base articles.  These articles correspond to possible questions or issues raised by callers; as a result, the call center employees have a reference to answer questions.  Additionally, this systemizes the format for recording issues for each call, since each topic has a standardized name.  This ensures topics and departments match across the list, allowing for efficient and accurate analysis.  Since this is the most recent data and is reflective of the city's current systems, this file was heavily relied upon for analysis on a departmental and topical basis.

@@

Exploration of calls by topic was a priority for the city.  Within each department, numerous topics exist.  Given the recent implementation of knowledge base articles (which can be seen in the most recent data), the attachment of topics to each call has been standardized.  This allows for a simplified, more accurate analysis.  Calls by topic were specifically analyzed by average duration and total call volume.
\par
When ranked by average duration, many obscure topics rose to the extremes.  Of the top 20 topics with the shortest duration, 19 had a total number of calls of less than three.  When viewed as a percentage of the total 202,500 call dataset, these calls are such a small proportion that it would give little actionable insight to the city.  A similar, though less extreme, situation occurred when viewing topics with longest duration.
\par
In light of this exploration, it was determined that the most important results for the city to see would include only topics that composed a significant proportion of total calls.  When ranking by duration, only topics with total volume above the first quartile of total call volume were included.  This yielded results that could be viewed as more relevant to management.  Topics with long durations seemed to be centralized around complaints and billing issues, and quick topics were mostly comprised of general information inquiries.  When analyzed in this manner, there is a 195 second difference between the longest and shortest topics.
\par
When topics were ranked in order of total call volume, the top results mostly represented the busiest departments identified earlier, such as Water Works and Solid Waste. The top topics included scheduling an extra trash pickup and water line shut-off deadlines.  Additionally, two of the most voluminous topics were ``Miscellaneous Trash Information" and ``Water Miscellaneous".  It could be inferred that these two topics serve as a catch-all for the call center.  By showing management these results, they may be able to discover the true nature of many of these calls, as well as impress the importance of accurate recording by call center operators.

@@

These charts all utilize various implementation of Bokeh tools, such as HoverTool, BoxZoom, Pan, and checkbox interactivity.  By incorporating these interactions into the dashboard, the result is a more dynamic, engaging product that is simple to interact with for employees outside of the technical fields.  An image of the final dashboard can be found at the end of this paper, while a live link can be accessed at \href{http://jdbul33.github.io/CallCenterDashboard.html}{http://jdbul33.github.io/CallCenterDashboard.html}.

@@
